---
title: "Obtaining and Cleaning a Data Set"
author: ""
output:
     html_document:
          toc: TRUE
---

*********

# Obtaining and Cleaning a Data Set

```{r include=FALSE}
library(knitr)
library(rmarkdown)
```

```{r global_options, eval=FALSE, include=FALSE}
knitr::opts_chunk$set(fig.width=12, fig.height=8, fig.path='Figs/',
                      echo=TRUE, warning=FALSE, message=FALSE)
```


## Libraries/Packages Required

The following libraries were used

| Package    | Version |
|------------|:--------|
| data.table | 1.11.4  |
| dplyr      | 0.7.6   |
| tidyr      | 0.8.1   |
| lubridate  | 1.7.4   |

If installed, they can be loaded as follows:

```{r libraries, message=FALSE, warning=FALSE, include=TRUE}
library(data.table)
library(dplyr)
library(tidyr)
library(lubridate)
```

## Obtaining & Extracting Data

###The Data

The data was accessed from:  
**URL:** <https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip>
**Time:** 2018-08-07 20:14:20 BST.  
**Data Format:** ZIP Archive

### Download of, and Extracting the Raw Data

If the data is not downloaded, the data is downloaded from the above URL. If the data has been previously downloaded and named as data.zip, the script prints a message and moves onto the next step.

```{r message=FALSE}
if (!file.exists("data.zip")) {
     message("Downloading data from 'https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", dest="data.zip", dest="data.zip'... Please wait")
     download.file(url = "https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", dest="data.zip", dest = "data.zip", method = "curl")
} else {
     message("Data already downloaded as data.zip")
}
```

If the data is not extracted, the data is extracted from data.zip. If the data has been previously extracted to to "UCI HAR Dataset", the script prints a message and moves onto the next step.

```{r message = FALSE}
if (file.exists("data.zip") & !file.exists("activity.csv")) {
     message("Extracting data... Please wait")
     unzip("data.zip")
} else {
     message("Data already extracted to ./activity.csv")
}
```

## Producing the Full Data Set
